Estimating initial state and quark-gluon plasma medium properties \\ using a hybrid model with nucleon substructure \\ calibrated to p-Pb and Pb-Pb collisions at \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV

### 2018-07-30 paper/substructure.tex
Writing summary

In this work, we constructed a multi-stage hybrid transport model for relativistic nuclear collisions and calibrated free parameters of the model using Bayesian parameter estimation.
Specifically, we focused on improving the model's description of small collision systems by investigating the effect of nucleon substructure on bulk observables in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV.

The analysis was performed using a parametric data-driven approach to account for theoretical uncertainty (variability) in the QGP initial conditions and hydrodynamic transport coefficients using a handful of equally uncertain model parametersi.
We marginalized over the theoretical uncertainty in the parameter using a handful of equally uncertain model parameters to vary the size and shape of the nucleon, the scaling behavior of initial QGP entropy deposition, the duration of pre-equilibrium dynamics, and the temperature dependence of the hydrodynamic shear and bulk viscosities.

We placed a conservative prior on the ``true'' value of each parameter, and calculated a global likelihood

varied each parameter within its design range to explored the full 15-dimensional parameter space of our model.

constructed an emulator to interpolate the predictions of our model at arbitrary regions of parameter space.

Bayesian inference was then used to compare the predictions of the model to experimental data from CMS and ALICE, in order to quantify the posterior probability that a given set of model parameters describes the experimental data

\medskip

>>>>

Relativistic heavy-ion collisions produce long-range multiparticle correlations which are commonly explained by the existence of hydrodynamic flow \note.
This narrative is evidenced by the global, self-consistent and highly non-trivial quantitative agreement of hydrodynamic models with a large number of heavy-ion bulk observables \note.
Naturally, such descriptions rely on the validity of hydrodynamic approximations, and these approximations begin to break down in the dilute particle limit where discrete particle degrees-of-freedom dominate and continuous field descriptions of the medium cease to make sense.
Thus unlike heavy-ion collisions, small system collisions were not expected to exhibit hydrodynamic behavior.

%Unlike heavy-ion collisions, small system collisions , were not expected
%
%Generally speaking, hydrodynamic flow signals were not expected to be observed
%
%
%Recent measurements measurements of multiparticle correlations in $p$-Pb collisions are surprisingly similar to those measured in Pb-Pb collisions.
%
%The hydrodynamic framework, however, is naturally expected to break down in the dilute limit where discrete particle degrees of freedom dominate and continuous field descriptions of the medium cease to make sense.
%It thus came somewhat as a surprise when high-multiplicity $p$-Pb collisions were found to produce long-range multiparticle correlations similar in magnitude to those observed in Pb-Pb collisions.
%The observation challenged preconceived notions of hydrodynamic applicability and suggested that even proton-lead collisions produce small droplets of QGP liquid which evolve hydrodynamically.
%
%The potential for a unified hydrodynamic description of small and large collision systems is of course very appealing, although it is by no means given.
%Multiple alternative explanations for the observed correlations have been proposed in the literature, and these mechanisms may contribute to the observed correlations in part or in full.
%
%Relativistic Au-Au and Pb-Pb collisions at RHIC and the LHC produce long-range multiparticle correlations which are commonly explained by the existence of hydrodynamic flow \note.
%Perhaps the strongest evidence for this claim has been the overwhelming success of hydrodynamic models in describing (and even predicting) a large number of soft-sector observables in relativistic heavy-ion collisions \note.
%
%The hydrodynamic framework, however, is naturally expected to break down in the dilute limit where discrete particle degrees of freedom dominate and continuous field descriptions of the medium cease to make sense.
%It thus came somewhat as a surprise when high-multiplicity $p$-Pb collisions were found to produce long-range multiparticle correlations similar in magnitude to those observed in Pb-Pb collisions.
%The observation challenged preconceived notions of hydrodynamic applicability and suggested that even proton-lead collisions produce small droplets of QGP liquid which evolve hydrodynamically.
%
%The potential for a unified hydrodynamic description of small and large collision systems is of course very appealing, although it is by no means given.
%Multiple alternative explanations for the observed correlations have been proposed in the literature, and these mechanisms may contribute to the observed correlations in part or in full.
%
\medskip
